---
title: "STAT 6550 HW6"
author: "Blain Morin"
date: "4/20/2021"
output: pdf_document
---

## 1. In Homework 2, we studied a time series of wind speeds (minus the gusts) recorded every twenty minutes at an airport in Fort Lauderdale, observed during the progression of Hurricane Wilma in October 2005. The data, on the square root scale, was provided in the R file for Homework 2. Start by removing the last 5 hours (15 time points) of data from the end of the time series. We will use these 15 values to assess forecasting performance in part (g). For parts (a) – (f) only use the remaining 123 values.


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(stargazer)
library(forecast)
library(kableExtra)
library(knitr)
library(gridExtra)

## Here is the dataset of wind speeds after a square root transformation.
winds <- c(2.90, 3.06, 3.62, 3.38, 3.43, 3.07, 3.13, 3.27, 3.35, 3.21, 3.26,
           3.51, 3.21, 2.83, 2.59, 2.62, 2.88, 2.75, 2.52, 2.68, 3.13, 3.22,
           3.60, 3.47, 3.30, 3.33, 3.38, 3.27, 3.32, 3.37, 4.04, 3.94, 3.74,
           3.77, 3.79, 3.80, 3.96, 3.87, 3.68, 4.01, 4.26, 4.16, 4.32, 3.93,
           3.86, 3.79, 3.89, 4.04, 3.46, 3.26, 3.17, 3.41, 3.67, 3.68, 3.88,
           3.84, 4.05, 4.10, 3.59, 3.75, 4.00, 3.67, 4.02, 4.35, 4.09, 4.03,
           3.99, 4.13, 4.15, 4.28, 4.51, 4.54, 4.44, 4.51, 4.50, 4.47, 4.52,
           4.43, 4.93, 5.15, 4.90, 5.09, 4.95, 5.24, 5.67, 5.69, 6.12, 6.30,
           6.25, 6.47, 6.63, 6.48, 6.41, 6.58, 6.97, 7.62, 8.03, 7.98, 7.67,
           7.26, 7.51, 7.65, 7.57, 7.50, 7.86, 7.85, 7.27, 6.77, 6.81, 6.83,
           6.64, 6.49, 6.27, 6.07, 5.96, 5.64, 5.52, 5.45, 5.37, 5.18, 4.89,
           4.53, 4.31, 4.37, 4.41, 4.54, 4.68, 4.80, 4.59, 4.42, 4.14, 4.00,
           4.17, 3.95, 3.97, 3.95, 3.89, 3.97)

## Create the 'hours' variable.
hours <- seq(from=0, by=1/3, length=length(winds))

```

## (a) Produce a properly labeled time series plot of the remaining 123 values.

```{r}

df = data.frame(winds = winds, hours = hours)

train = df[1:123,]
test = df[124:138, ]


```

```{r, fig.height=4}

train %>%
  ggplot(aes(x = hours, y = winds)) +
  geom_line(size = 1.5) +
  ylab("Wind Speed (sqrt scale)") +
  xlab("Hour") +
  ggtitle("Time Series Plot for First 123 Values") +
  theme_bw()

```

\newpage

## (b) Follow the steps to carry out the Dickey-Fuller test on the square root wind speeds, as outlined in the “Modeling ARIMA and SARIMA processes” class notes. Show all your work and write down your conclusion in words.

```{r, results='asis', echo = TRUE}

### Create data frame
### y is the first difference
### x is the wind speed at the previous timepoint
df.test = data.frame(y = diff(train$winds),
                     x = df$winds[1:122],
                     hours = hours[1:122])


### Linear Model
model = lm(y ~ x, data = df.test)
sum.model = summary(model)

### Make a table with results
stargazer(model, header = FALSE, title = "Dickey-Fuller Regression Results")

```

The Dickey-Fuller statistic is $-.019/.015 = -1.277$. Using the Dickey-Fuller distribution critical value, $C_{\alpha = .05} = -2.86$, we fail to reject $H_0$. Thus, there is not enough evidence to suggest that the process is stationary.

\newpage

## (c) Based on a time series plot of the differenced square root wind series discuss whether or not stationarity is a reasonable assumption for the differenced series.

```{r, fig.height=4}

df.test %>%
  ggplot(aes(x = hours, y = y)) +
  geom_line(size = 1.5) +
  ylab("Differenced sqrt Wind Speed") +
  xlab("Hour") +
  ggtitle("Differenced Series Plot") +
  theme_bw()

```


Visually, the differenced series looks mostly stationary. We see it is mean 0 and has relatively constant variance. There is a bit of a dip after the 35 hour mark, which is maybe a bit concerning.

## (d) Examining the first fifty lags of the sample ACF and sample PACF of the differenced series, explain why an ARIMA(1,1,0) model may be reasonable for the undifferenced square root wind series.

```{r}

a = ggAcf(df.test$y, lag.max = 50) +
  ggtitle("ACF Plot") +
  theme_bw()

b = ggPacf(df.test$y, lag.max = 50) +
  ggtitle("PACF Plot") +
  theme_bw()

grid.arrange(a, b, nrow = 2)

```

On the PACF plot of the differenced series, we see that the first lag is outside of the interval. We see a similar pattern in the ACf plot as well. An ARIMA(1,0,0) model would be a good candidate for the differenced series, thus a an ARIMA(1,1,0) model would be a reasonable choice for the undifferenced series (they are equivalent).

\newpage 

## (e) Now compare the following time series models fit to the square root wind series:

* (i) ARIMA(0,1,0) (a random walk model) without an intercept
* (ii) ARIMA(1,1,0) without an intercept
* (iii) A subset ARIMA(8,1,0) model without an intercept in which only $\phi_1$ and $\phi_8$ are non-zero.

```{r}

plot.residuals <- function (x, y=NULL, lag.max=NULL, mean.line=TRUE,
			    acf.ylim=c(-0.25,1), mfrow=c(2,2),
                            lags=NULL, ...)
{
  ## we want a 'mfrow[1] x mfrow[2]' display for the
  ## plots, only if mfrow is not NULL
  if (!is.null(mfrow))
    par(mfrow=mfrow)
  
  ## if we do not supply a time variable.
  if (is.null(y))
  {
    y <- x
    x <- seq(length(y))
  } else { ## make sure we get no trouble from time series objects!
    x <- as.numeric(x)
    y <- as.numeric(y)
  }

  if (is.null(lag.max)) {
    lag.max <-  floor(10 * log10(length(x)))
  }

  ## Produce a time series plot of the residual,
  ## adding a line at 'y'=0 if 'mean.line' is TRUE.  
  plot(x, y, type="l", ...)
  if (mean.line) abline(h=0, lty=2)

  ## Produce a normal Q-Q plot.
  qqnorm(y, main="")
  qqline(y)

  ## Plot the sample ACF and PACF
  if (is.null(lags)) {
    acf(y, main="", lag.max=lag.max, xlim=c(0, lag.max), ylim=acf.ylim,
        ylab="sample ACF")

    pacf(y, main="", lag.max=lag.max, xlim=c(0, lag.max), ylim=acf.ylim,
         ylab="sample PACF")
  }  
  else {
    acf(y, main="", lag.max=lag.max, xlim=c(0, lag.max), ylim=acf.ylim,
        ylab="sample ACF", xaxt="n")
    axis(side=1, at=lags)

    pacf(y, main="", lag.max=lag.max, xlim=c(0, lag.max), ylim=acf.ylim,
         ylab="sample PACF", xaxt="n")
    axis(side=1, at=lags)
  }

  ## Finally carry out the Ljung-Box test
  ## We will use the same number of lags as used in the ACF and PACF plot.
  Box.test(y, lag.max, type="Ljung-Box")
}

```

### ARIMA(0,1,0)

```{r, echo = TRUE}

arima010 = arima(train$winds, order = c(0,1,0), include.mean = FALSE)

resids010 = resid(arima010)

arima010

### Function From Class Notes
plot.residuals(resids010)

```

The random walk diagnostics look pretty good. The residuals are mean zero and have mostly constant variance. We see from the QQ plot that there are some slight deviations from normality in the tails of the residual distribution. No lag bars are outside of the interval in the ACF or PACF plots. The Box-Ljung test is not significant, meaning that there is not enough evidence to say that the residuals are not IID. 


\newpage

### ARIMA(1,1,0)

```{r, echo = TRUE}

arima110 = arima(train$winds, order = c(1,1,0), include.mean = FALSE)

resids110 = resid(arima110)

arima110

plot.residuals(resids110)

```

The diagnostic plots seem improved for the ARIMA(1,1,0) model compared to the random walk model. We can visually tell that the AR1 95\% confidence interval does not contain 0, meaning that it is significant. The points on the QQ plot are near the line, indicating that the residuals are normally distributed. There is no evidence from the ACF, PACF, or Box-Ljung test rhat are residuals are not IID. 

\newpage

### Subset ARIMA(8,1,0)

```{r, echo = TRUE}

arima810 = arima(train$winds, order =c(8,1,0), include.mean = FALSE,
                 fixed = c(NA, rep(0, 6), NA))

resids810 = resid(arima810)

arima810

plot.residuals(resids810)

```

The diagnostic plots are also good for the subset model. We see visually that both the AR1 and AR8 components are statistically significant (there confidence intervals do not contain 0). Again, the residuals are mean 0 with relatively constant variance. The points in the QQ plot are near the line, indicating that the residuals are normally distributed. There is nothing in the ACF, PACF, or Box-Ljung test to suggest that the residuals are not IID.

\newpage

## (f) Now compare the AICC values for the above three models. Which model “fits better”, with respect to this criterion?


```{r}
AIC.to.AICC <- function (aic, n, npars) {
  aic - 2 * npars * ( 1 - n/(n-1-npars))
}

n = length(train$winds)

```

```{r, echo = TRUE}

### Function from Class Notes
AIC.to.AICC(arima010$aic, n , 1)

AIC.to.AICC(arima110$aic, n, 2)

AIC.to.AICC(arima810$aic, n, 3)

```

According to the AICC (lower is better), the subset ARIMA(8,1,0) model has the best fit.

